---
title: "oGraph"
author: "Xin He"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
%\VignetteIndexEntry{oGraph}
%\VignetteEngine{knitr::rmarkdown}
\usepackage[utf8]{inputenc}
---
  
# Overview
oGraph is a package for ontology representation and visualization. 

# Quick start guide
This section describes a simple working session using **oGraph**. 

A typical session can be divided into two steps:
  
1.  *Data preparation* build the oGraph object which has the ontology store in an underlying igraph object.

2.  *Running analysis* treeview, subtree, find parent/child terms...


Here are example vignettes:

```{r 1,message=FALSE}
library('ograph')
ograph::initOGraph('HDO')
```

```{r 2,message=FALSE}
ontGraph<-new("ontGraph",ontology='HDO')
print(ontGraph)
g=ontGraph@graph
root=findRoot(graph=g)
leaves=findLeafNode(graph=g)
treeplot(g,label=0)

nodes=shortest_path_to_root(graph=g,node='DOID:10652')
g2=set.node.attribute(g,attr_name='color',attr_value='red',nodes=nodes)
treeplot(g2,label=0)
```
</center>


```{r}
subg<-subGraphByLevel(graph=g,l=3)
treeplot(subg,label=1)
```
  
```{r}
terms=c('DOID:10652','DOID:14330')
list.vertex.attributes(graph=g)
get.node.attribute(g,attr='def',nodes=terms)
subg<-subGraphByNodes(graph=g,nodes=terms)
treeplot(subg,label=1)
```


```{r}
subg=mapGene2Graph(graph=subg,file=system.file("extdata","human_gene2HDO_o", package ="ograph"))
treeplot(subg,label=1,show.genes=TRUE)
```


```{r}
term='DOID:150'
get.node.attribute(g,attr='def',nodes=term)
peekNode(g,node=term)

findParentNodes(graph=g,node=term)
findChildrenNodes(graph=g,node=term)

cs=findAllChildrenNodes(g,term)
g.cs=subGraphByNodes(graph=g,nodes=cs,include_path_to_root=TRUE)
treeplot(g.cs,label=0)

ps=findAllParentNodes(g,term)
g.ps=subGraphByNodes(graph=g,nodes=ps,include_path_to_root=TRUE)
treeplot(g.ps,label=1)
```


```{r}
clip_neuro=searchDescription4Keyword(graph=g,keys=c('neuro'))
names=searchName4Keyword(g,keys=c('neuro'))
get.node.attribute(g,attr='def',names)

clip_brain=searchDescription4Keyword(graph=g,keys=c('brain'))
g.clip=subGraphByNodes(g,nodes=unique(c(clip_brain,clip_neuro)))
g.clip=set.node.attribute(graph=g.clip,attr_name='color',attr_value='green',nodes=clip_neuro)
g.clip=set.node.attribute(graph=g.clip,attr_name='color',attr_value='blue',nodes=clip_brain)
g.clip=set.node.attribute(graph=g.clip,attr_name='color',attr_value='red',nodes=intersect(clip_brain,clip_neuro))

treeplot(g.clip,label=0)
```
